\section{Introduction}
\label{sec:intro}
% trungnt
% Remote terminal access remains a fundamental interface for operating servers, cloud instances, edge devices, and experimental testbeds. While traditionally driven by human operators, recent automation workflows increasingly exercise remote shells through repeated command execution, output parsing, and interactive terminal updates. These automated high-frequency interactions are highly sensitive to transport-level bottlenecks. SSHv2 remains the de facto protocol for remote shell access, but its TCP transport suffers from head-of-line (HoL) blocking and binds sessions to fixed network addresses, causing severe stalls during automated command execution under network impairments \cite{Scharf2007}. Modern alternatives utilizing UDP (Mosh \cite{Winstein2012}) or HTTP/3 and QUIC (SSH3 \cite{michel2023}) address these legacy weaknesses, yet their design focus on human typing and network roaming leaves their reliability for deterministic, burst-driven agent workloads unclear. Therefore, this paper asks: \textbf{Which SSH Connection Protocol can meet the strict latency and completeness demands of closed-loop terminal agents?}

% Evaluating the performance of terminal protocols for autonomous agents is inherently challenging because current methodologies isolate either the human experience or the AI logic, but rarely their intersection. In the networking domain, foundational evaluations of protocols like Mosh and SSH3 heavily prioritize human-centric metrics \cite{michel2023}. They measure single-character echo latency and connection resilience during IP roaming, assuming input rates of just a few words per minute. Conversely, recent AI research introduces sophisticated terminal environments to evaluate the reasoning capabilities of AI. However, these AI benchmarks treat the underlying network as a frictionless, zero-latency pipe, focusing exclusively on end-to-end task success \cite{Liu2024}. This dichotomy creates a critical evaluation gap. When a transport protocol stalls due to packet loss or drops terminal output during a high-frequency automated burst, the agent’s closed-loop observation process is fundamentally disrupted. Consequently, existing literature lacks a targeted methodology to measure how different transport mechanisms compromise the completeness and timing of AI-driven interactions. To address this gap, this paper does not evaluate end-to-end AI task success. Instead, we introduce a deterministic benchmark that models an AI-like interactive client. By isolating transport-level behavior, we evaluate the latency and completeness of the feedback channel that closed-loop systems experience over these protocols. 

% We design a workload suite that captures three distinct interaction patterns to AI clients: short-output command loops, large-output data streams, and complex single-pane and multi-pane interactive editing sessions. We deploy these workloads across SSHv2, SSH3, and Mosh to quantify per-workload command-response latency and output completeness. To ensure robust validation, we evaluate these metrics across two distinct testing topologies: a controlled testbed featuring three defined network impairment profiles, and a real-world VPN deployment. The main contributions of this paper are summarized as follows:

% % The goal is to isolate the transport-level behavior that such closed-loop clients would experience before adding model reasoning, planning, or task-specific policy decisions.
% % trungnt


% \begin{itemize}
%     \item We conduct a deterministic measurement study evaluating SSHv2, SSH3, and Mosh across real-world VPN and controlled network impairments, specifically isolating transport-level bottlenecks for AI-driven terminal workloads.
    
%     \item We show that SSH3's QUIC-based architecture is optimal for agile, short-lived agent interactions. It achieves the lowest session setup latency and provides the fastest command-response loop for short terminal outputs.
    
%     \item We reveal a tradeoff between human-visible responsiveness and programmatic reliability for large-output commands. While Mosh excels in interactive keystroke latency, its state-synchronization model drops over 98\% of expected bytes, breaking the AI observation loop. Conversely, legacy SSHv2 strictly preserves 100\% byte completeness and completes large data streams significantly faster than SSH3.
% \end{itemize}

% We discuss background and related work in Section~\ref{sec:background}, detail our methodology in Section~\ref{sec:methodology}, and present our evaluation in Section~\ref{sec:evaluation}, before concluding in Section~\ref{sec:conclusion}.


Remote terminal access remains a fundamental interface for
operating servers, cloud instances, edge devices, and experimental testbeds. Although such access has traditionally been driven by human operators, it is increasingly exercised by automated scripts, programmatic clients, and AI-agent-oriented workflows. These clients interact with remote systems through repeated command execution, output parsing, and terminal updates. In such settings, the remote shell becomes part of an action--observation loop: a command is issued, the terminal response is observed, and the next action depends on the received output. As a result, transport latency, tail delay, and output completeness can directly affect the
quality of the terminal feedback channel.

SSHv2~\cite{Ylonen2006RFC4253} remains the de facto protocol for secure remote shell access.
However, its TCP-based transport inherits ordered byte-stream
delivery and can suffer from head-of-line (HoL) blocking under
packet loss or delay variation~\cite{Scharf2007}. Moreover, conventional SSHv2 sessions are tied to network addresses, making them fragile under mobility or network changes. Modern alternatives explore different design points. Mosh~\cite{Winstein2012} uses a UDP-based state synchronization protocol to improve visible responsiveness and support roaming, while SSH3~\cite{michel2023} redesigns secure shell access over HTTP/3 and QUIC. These mechanisms are promising for interactive use, but their behavior under automated, burst-driven, and output-sensitive terminal workloads remains insufficiently understood.

A key challenge is that existing evaluation methodologies usually focus on only one side of the problem. Networking studies of remote terminal protocols often emphasize human-perceived responsiveness, such as keystroke echo latency, screen freshness, or roaming resilience. In contrast, AI-agent and automation benchmarks usually evaluate task completion or reasoning quality while treating the remote shell and the network path as an ideal execution substrate. This separation leaves an important gap: we do not yet know how different remote control protocols behave when the terminal workload
resembles the interaction pattern commonly induced by automated or
AI-agent-oriented clients.

To address this gap, we focus on the terminal feedback channel rather than evaluating end-to-end agent task success. Our goal is to characterize the transport-level behavior of remote shell protocols under agent-like terminal interactions, independently of model reasoning time, task planning, and application-specific agent logic. We therefore use a deterministic benchmark client to emulate representative terminal feedback patterns, including short command loops, large-output data streams, and interactive editing under both single-pane and multi-pane terminal activity.

We compare SSHv2, SSH3, and Mosh under both controlled network impairments and a VPN-based Internet path. The evaluation measures session setup latency, command/output delivery latency, keystroke responsiveness, and output completeness. These metrics capture different requirements of agent-like terminal interactions: fast startup for short remote-control loops, low response latency for small commands, byte-complete delivery for large outputs, and visible responsiveness during interactive editing. The main contributions of this paper are summarized as follows:
\begin{itemize}
    \item We present a deterministic measurement study of SSHv2, SSH3, and Mosh under agent-like interactive terminal workloads across controlled network impairments and a VPN-based Internet path.

    \item We design a scripted workload suite that captures common terminal interaction patterns induced by automated and AI-agent-oriented clients, including short command loops, large-output streams, and interactive editing.

    \item We quantify the trade-off between visible responsiveness and programmatic reliability by jointly measuring latency and output completeness across different protocol designs.

    \item We derive workload-dependent protocol selection guidelines: SSH3 is preferable for fast setup and short command-response loops, SSHv2 remains stronger for byte-complete large-output delivery, and Mosh is best suited for human-visible interactive responsiveness but not for output-sensitive automated parsing.
\end{itemize}

The remainder of this paper is organized as follows. Section~II discusses background and related work. Section~III presents the benchmark design, workloads, metrics, and experimental setup. Section~IV reports and analyzes the evaluation results. Finally, Section~V concludes the paper.
