\section{Background and Related Work}
\label{sec:background}
% Remote shell protocols expose different notions of responsiveness to an Agent-like terminal client. A session can be considered effective because it becomes usable quickly, because short command responses arrive with low latency, because typed characters become visible despite packet loss, or because command output is delivered as a complete byte stream. These goals are not equivalent. A protocol that minimizes visible delay may discard intermediate terminal states, while a protocol that preserves every byte may delay feedback until missing data has been retransmitted. This section summarizes the protocol mechanisms behind these trade-offs and relates them to prior measurement studies.

% SSHv2 \cite{Ylonen2006RFC4251} is the de facto mechanism for secure remote terminal access. It runs over TCP and provides ordered, reliable delivery for encrypted SSH channels \cite{Bellare2002}. This byte-stream semantics is useful when the client must receive complete command output for programmatic parsing, but it also inherits TCP HoL blocking: a lost segment can delay later data until retransmission completes \cite{Scharf2007}. For a closed-loop terminal client, this can appear as delayed command completion, delayed output availability, or delayed visible feedback. SSHv2 setup also includes a TCP handshake followed by SSH key exchange and authentication, so the setup cost grows with round-trip time (RTT) \cite{Ylonen2006RFC4253}. These properties make SSHv2 a natural reference point for byte-complete delivery, especially in workloads that stream large outputs through a single shell channel.

% SSH3 redesigns SSH over HTTP/3 and QUIC \cite{michel2023}. Its main architectural promise is lower setup latency: QUIC integrates transport and cryptographic negotiation, enabling 1-RTT setup and 0-RTT resumption for eligible connections \cite{Iyengar2021}. QUIC also removes HoL blocking across independent streams. For terminal workloads, however, the benefit depends on how shell traffic is mapped onto streams. If a workload consists of independent channels, QUIC stream multiplexing can prevent loss in one stream from blocking unrelated streams. If a workload produces one ordered terminal output flow, such as a large command printed through a single interactive shell, the stream still has to be completed in order before byte-complete output is available to the client. Consequently, SSH3 should be evaluated separately for session setup, short command-response loops, and large single-stream outputs rather than treated as uniformly faster than SSHv2.

% Mosh makes the opposite trade-off \cite{Winstein2012}. It uses SSH only to authenticate and bootstrap the session, then switches to a UDP channel carrying terminal state updates through the State Synchronization Protocol (SSP). SSP treats the terminal as a replicated screen state and sends compact updates for the latest visible terminal contents. This design avoids waiting for every intermediate byte and makes lost datagrams superseded by newer state. Mosh further improves perceived responsiveness through predictive local echo. The cost is completeness: output that scrolls past between screen updates may never be delivered to the client. This is acceptable for many human editing sessions, where the latest visible screen is often sufficient, but it is risky for automated clients that parse command output as data. Therefore, Mosh latency measurements must be interpreted together with output completeness; a fast observable completion is not necessarily a byte-complete completion.

% Mosh is designed for interactive use on mobile networks. Its state synchronization and predictive local response help reduce keyboard operation latency compared to SSHv2. QUIC studies provide useful context for SSH3, although their results are controversial. When QUIC is deployed on a large scale, reports show improvements in latency and user experience, while controlled measurements suggest that benefits depend on the deployment method, workload, and network path. Recent throughput studies also identify thread concurrency and packet processing in the user space as critical factors in high-volume transmission performance. These results suggest that the performance of the remote shell should be tested separately for short control messages, large outputs, and interactive screen updates.

% The agent performance tests drive these terminal workloads but do not evaluate the underlying transport mechanisms. SWE-agent interacts with the system via shell commands, file manipulation, and terminal observables, while Terminal-Bench evaluates automated tasks in a command line environment. Other performance tests also focus on task completion and often treat the shell as a reliable local interface. Therefore, they do not capture how network latency or incomplete output displayed to the client affects the agent's response loop.

% This raises an open question: How do SSHv2, SSH3, and Mosh perform across different terminal response models? Existing studies do not provide a controlled comparison on the client side of session setup, small/large command responses, output completeness, and interaction latency in both controlled simulated network scenarios and Internet paths. We address this gap by using a simulated terminal workload, separating transport behavior from model reasoning and task-specific decisions.

%trungnt
Remote shell protocols expose different notions of responsiveness to an
agent-like terminal client. A session can be considered effective because it becomes usable quickly, because short command responses arrive with low
latency, because typed characters become visible despite network impairment, or because command output is delivered as a complete byte stream. These goals are not equivalent. A protocol that minimizes visible delay may discard intermediate terminal states, while a protocol that preserves every byte may delay feedback until missing data has been retransmitted. This section summarizes the protocol mechanisms behind these trade-offs and relates them to prior measurement studies.

SSHv2~\cite{Ylonen2006RFC4251} is the de facto mechanism for secure remote
terminal access. It runs over TCP and provides ordered, reliable delivery for encrypted SSH channels~\cite{Bellare2002}. This byte-stream semantics is useful when the client must receive complete command output for programmatic parsing, but it also inherits TCP head-of-line (HoL) blocking: a lost segment can delay later data until retransmission completes~\cite{Scharf2007}. For a closed-loop terminal client, this can appear as delayed command completion, delayed output availability, or delayed visible feedback. SSHv2 setup also includes a TCP handshake followed by SSH key exchange and authentication, so the setup cost grows with round-trip time (RTT)~\cite{Ylonen2006RFC4253}. These properties make SSHv2 a natural reference point for byte-complete delivery, especially in workloads that stream large outputs through a single shell channel.

SSH3 redesigns SSH over HTTP/3 and QUIC~\cite{michel2023}. Its main
architectural promise is lower setup latency: QUIC integrates transport and cryptographic negotiation, enabling 1-RTT setup and 0-RTT resumption for eligible connections~\cite{Iyengar2021}. QUIC also removes HoL blocking across independent streams. For terminal workloads, however, the benefit depends on how shell traffic is mapped onto QUIC streams. If a workload uses independent channels, stream multiplexing can prevent loss in one stream from blocking unrelated streams. If a workload produces one ordered terminal output flow, such as a large command printed through a single interactive shell, bytes within that stream still have to be delivered in order before byte-complete output is available to the client. Consequently, SSH3 should be evaluated separately for session setup, short command-response loops, and large single-stream outputs rather than treated as uniformly faster than SSHv2.

Mosh makes a different trade-off~\cite{Winstein2012}. It uses SSH only to
authenticate and bootstrap the session, then switches to a UDP channel
carrying terminal state updates through the State Synchronization Protocol
(SSP). SSP treats the terminal as a replicated screen state and sends compact updates for the latest visible terminal contents. This design avoids waiting for every intermediate byte and allows lost datagrams to be superseded by newer state. Mosh further improves perceived responsiveness through predictive local echo. The cost is completeness: output that scrolls past between screen updates may never be delivered to the client. This is acceptable for many human editing sessions, where the latest visible screen is often sufficient, but it is risky for automated clients that parse command output as data. Therefore, Mosh latency measurements must be interpreted together with output completeness; a fast observable completion is not necessarily a byte-complete completion.

Prior measurement studies provide useful context but do not directly answer the workload-specific question addressed in this paper. Mosh was originally evaluated for interactive use on mobile networks, where state synchronization and predictive local echo reduce keyboard response latency compared with SSHv2~\cite{Winstein2012}. QUIC studies report performance benefits in some large-scale deployments, but controlled measurements show that these benefits depend on workload, implementation, deployment method, and network path~\cite{Langley2017,Kakhki2017}. Recent throughput studies further show that QUIC performance can be affected by user-space packet processing, thread concurrency, and stream usage~\cite{Konig2025}. These results suggest that remote shell protocols should be evaluated separately for short control messages, large terminal outputs, and interactive screen updates.

Agent benchmarks motivate these terminal interaction patterns but usually do not evaluate the underlying transport mechanisms. SWE-agent interacts with software environments through shell commands, file manipulation, and terminal observations~\cite{Yang2024}. Terminal-Bench evaluates automated task execution in command-line environments~\cite{Merrill2026}. Other agent benchmarks similarly focus on task completion and reasoning capability, often treating the shell as a reliable local interface. As a result, they do not capture how network latency, tail delay, or incomplete terminal output affects the feedback channel observed by an automated client.

This raises an open question: how do SSHv2, SSH3, and Mosh perform across
different terminal response models? Existing studies do not provide a
controlled client-observed comparison of session setup latency, small-command response latency, large-output delivery, output completeness, and interactive keystroke latency across both controlled network impairment scenarios and VPN-based Internet paths. We address this gap using scripted terminal workloads that separate transport behavior from model reasoning, task-specific planning, and application-level agent logic.
