\section{Performance Evaluation}
\label{sec:evaluation}

\subsection{Session Establishment Latency}
\label{subsec:session-setup}

Table~\ref{tab:session-setup} reports the session establishment latency across all three protocols and network scenarios. Session setup time is measured from process spawn to the first shell prompt, captured before any background tasks or workload iterations begin, isolating pure transport and authentication overhead.

\begin{table}[t]
\centering
\caption{Session establishment latency (ms). $N=18$ per cell.}
\label{tab:session-setup}
\begin{tabular}{llrrrrr}
\hline
\textbf{Scenario} & \textbf{Protocol} & \textbf{Mean} & \textbf{Median} & \textbf{Std} & \textbf{P95} & \textbf{Max} \\
\hline
\multirow{3}{*}{Low}
 & SSHv2 & 811 & 816 & 49  & 879  & 879  \\
 & SSH3  & 362 & 376 & 50  & 430  & 433  \\
 & Mosh  & 1166 & 1171 & 36 & 1205 & 1219 \\
\hline
\multirow{3}{*}{Medium}
 & SSHv2 & 1796 & 1772 & 137 & 2080 & 2105 \\
 & SSH3  & 645  & 637  & 88  & 796  & 918  \\
 & Mosh  & 2366 & 2258 & 283 & 2917 & 3282 \\
\hline
\multirow{3}{*}{High}
 & SSHv2 & 3565 & 3527 & 706 & 4749 & 5184 \\
 & SSH3  & 1042 & 954  & 202 & 1459 & 1611 \\
 & Mosh  & 3856 & 3813 & 353 & 4523 & 4705 \\
\hline
\end{tabular}
\end{table}

SSH3 achieves the fastest session setup across all scenarios, completing 2.2--3.4$\times$ faster than SSHv2 and 3.2--3.7$\times$ faster than Mosh. Under the Low scenario (RTT $\approx$ 20\,ms), SSH3 establishes a session in 362\,ms (median 376\,ms), compared to 811\,ms for SSHv2 and 1166\,ms for Mosh.

This difference stems from the handshake design. SSH3 uses QUIC's 1-RTT handshake, which combines transport and cryptographic setup in a single round trip. SSHv2 requires a TCP three-way handshake followed by key exchange and authentication (2--3 RTTs). Mosh incurs the highest cost because it first completes a full SSH session to bootstrap its UDP-based SSP channel, adding an extra round trip on top of the SSH handshake.

As impairment increases, the absolute gap widens. Under the High scenario (RTT $\approx$ 200\,ms, 3\% loss), SSHv2 requires 3.6\,s and Mosh 3.9\,s on average, while SSH3 completes in 1.0\,s. Variability also grows disproportionately for TCP-based protocols: SSHv2's standard deviation reaches 706\,ms versus 202\,ms for SSH3, reflecting TCP's sensitivity to packet loss during multi-round-trip handshakes.

\subsection{Command Completion Latency (W1)}
\label{subsec:w1-results}

Table~\ref{tab:w1-lightweight} summarizes the command completion latency for lightweight commands (\texttt{ls}, \texttt{df~-h}, \texttt{grep}, \texttt{cat~/proc/meminfo}), while Table~\ref{tab:w1-heavy} reports results for heavy-output commands (\texttt{ps~aux}, \texttt{find~/usr~-maxdepth~3}). Each cell aggregates $N=300$ non-warmup samples (3 trials $\times$ 100 iterations) for the LAN scenarios.

\begin{table}[t]
\centering
\caption{W1 latency for lightweight commands (ms). Values are aggregated across four commands ($N=1200$ for SSH/SSH3, varies for Mosh due to failures).}
\label{tab:w1-lightweight}
\begin{tabular}{llrrrr}
\hline
\textbf{Scenario} & \textbf{Protocol} & \textbf{N} & \textbf{Mean} & \textbf{Median} & \textbf{P95} \\
\hline
\multirow{3}{*}{Low}
 & SSHv2 & 1200 & 50.5 & 50.5 & 50.8 \\
 & SSH3  & 1200 & 50.5 & 50.5 & 50.8 \\
 & Mosh  & 456  & 50.6 & 50.5 & 50.8 \\
\hline
\multirow{3}{*}{Medium}
 & SSHv2 & 1200 & 137.9 & 94.5  & 263.3 \\
 & SSH3  & 1200 & 133.8 & 155.8 & 248.7 \\
 & Mosh  & 515  & 122.9 & 111.9 & 178.6 \\
\hline
\multirow{3}{*}{High}
 & SSHv2 & 1200 & 262.7 & 205.9 & 572.5 \\
 & SSH3  & 1200 & 303.8 & 276.4 & 558.2 \\
 & Mosh  & 383  & 242.0 & 243.4 & 441.8 \\
\hline
\end{tabular}
\end{table}

\begin{table}[t]
\centering
\caption{W1 latency for heavy-output commands (ms). \texttt{ps~aux} ($\sim$4\,KB) and \texttt{find~/usr~-maxdepth~3} ($\sim$12\,KB).}
\label{tab:w1-heavy}
\begin{tabular}{llrrrrr}
\hline
\textbf{Scenario} & \textbf{Protocol} & \textbf{N} & \textbf{Mean} & \textbf{Median} & \textbf{P95} & \textbf{Stdev} \\
\hline
\multirow{3}{*}{Low}
 & SSHv2 & 600 & 110.0 & 71.5  & 186.0 & 57.3 \\
 & SSH3  & 600 & 133.0 & 62.3  & 241.3 & 81.8 \\
 & Mosh  & 600 & 168.0 & 58.8  & 297.1 & 119.4 \\
\hline
\multirow{3}{*}{Medium}
 & SSHv2 & 600 & 1448.9 & 682.3  & 3561.4  & 1235.3 \\
 & SSH3  & 600 & 2606.3 & 551.3  & 5942.3  & 2427.8 \\
 & Mosh  & 348 & 359.7  & 381.2  & 488.4   & 123.8 \\
\hline
\multirow{3}{*}{High}
 & SSHv2 & 600 & 3709.8  & 1662.7 & 9570.0  & 3452.1 \\
 & SSH3  & 600 & 7072.0  & 1520.7 & 15411.7 & 6517.9 \\
 & Mosh  & 53  & 431.8   & 463.4  & 794.9   & 189.7 \\
\hline
\end{tabular}
\end{table}

\subsubsection{Lightweight commands}
For commands producing small output ($<$1\,KB), all three protocols perform similarly under the Low scenario ($\sim$50.5\,ms), as the network overhead is negligible on a clean LAN link. Under impairment, the protocols diverge moderately. In the High scenario, Mosh achieves the lowest mean latency (242\,ms), followed by SSHv2 (263\,ms) and SSH3 (304\,ms). Mosh's advantage here comes from SSP's frame-based delivery, which avoids retransmission delays for small payloads.

SSH3's slightly higher latency under the High scenario (304\,ms vs.\ 263\,ms for SSHv2) is notable. This may reflect overhead in the SSH3 implementation (v0.1.7), which is still experimental, or QUIC's more conservative congestion control behavior on lossy links compared to TCP's established loss recovery.

\subsubsection{Heavy-output commands}
For commands producing larger output, the protocol differences become dramatic and reveal a critical trade-off. Under the High scenario, SSHv2 averages 3.7\,s and SSH3 averages 7.1\,s, while Mosh records only 432\,ms.

SSH3's poor performance for bulk transfer under impairment is unexpected. At 7.1\,s (High), SSH3 is 1.9$\times$ slower than SSHv2 (3.7\,s). This likely reflects the immaturity of SSH3's QUIC implementation: its congestion controller may react more conservatively to loss than TCP's well-tuned algorithms, and the per-stream framing adds overhead for large sequential transfers.

Mosh's low latency (432\,ms in High) is misleading. As shown in Section~\ref{subsec:output-completeness}, Mosh delivers only 2.5\% of the expected output for these commands. SSP transmits terminal frame states rather than raw byte streams; when output arrives faster than the display refresh interval, intermediate frames are discarded. The benchmark measures time-to-prompt, not time-to-complete-output.

\subsubsection{Failure rates}
SSHv2 and SSH3 achieve perfect reliability (0 failures across all scenarios and commands). Mosh exhibits significant failures, particularly for commands with larger output: only 53 successful samples out of 300 attempts for heavy commands in the High scenario (82\% failure rate). Even for lightweight commands, Mosh's effective sample count drops to 383--515 (vs.\ 1200 for SSH/SSH3), indicating 57--68\% of sessions failed to complete.

\subsection{Output Completeness}
\label{subsec:output-completeness}

A key metric for AI agent workloads is whether the client receives the complete command output. Table~\ref{tab:recv-pct} reports the mean percentage of expected output bytes received by the client (\texttt{received\_pct}), computed by comparing the captured output length against a reference collected via direct SSH execution.

\begin{table}[t]
\centering
\caption{Output completeness (\% of expected bytes received). Higher is better.}
\label{tab:recv-pct}
\begin{tabular}{llrr}
\hline
\textbf{Scenario} & \textbf{Protocol} & \textbf{Lightweight} & \textbf{Heavy-output} \\
\hline
\multirow{3}{*}{Low}
 & SSHv2 & 99.2 & 96.3 \\
 & SSH3  & 99.2 & 96.6 \\
 & Mosh  & 92.2 & 6.3  \\
\hline
\multirow{3}{*}{Medium}
 & SSHv2 & 99.2 & 98.1 \\
 & SSH3  & 99.0 & 97.0 \\
 & Mosh  & 90.5 & 2.1  \\
\hline
\multirow{3}{*}{High}
 & SSHv2 & 99.2 & 96.4 \\
 & SSH3  & 99.2 & 95.4 \\
 & Mosh  & 89.1 & 2.5  \\
\hline
\end{tabular}
\end{table}

SSHv2 and SSH3 both maintain high output completeness across all scenarios ($\geq$95\%), confirming that their byte-stream transport reliably delivers the full command output regardless of network conditions.

Mosh presents a fundamentally different picture. For heavy-output commands, Mosh delivers only 2--6\% of the expected output across all scenarios. Even for lightweight commands producing less than 1\,KB, Mosh's completeness drops to 89--92\%. This is not a failure mode but an inherent property of SSP: it synchronizes the terminal display state rather than the output byte stream. When multiple screen updates arrive within a single frame interval ($\sim$60\,ms), only the final state is transmitted.

For interactive human users, this trade-off is acceptable---the user sees the final result. For AI agents that must parse intermediate output (e.g., extracting specific lines from \texttt{ps~aux} or parsing directory listings from \texttt{find}), Mosh's frame-skipping behavior means the agent operates on incomplete data. This can lead to incorrect parsing, missed information, and cascading errors in the agent's decision loop.

\subsection{Continuous Monitoring Latency (W2)}
\label{subsec:w2-results}

% TODO

\subsection{Interactive Keystroke Latency (W3)}
\label{subsec:w3-results}

% TODO

\subsection{Large Output Delivery (W4)}
\label{subsec:w4-results}

% TODO

\subsection{Key Takeaways}
\label{subsec:takeaways}

% TODO
